---
title: "Profit Model"
author: "Rita Liu"
output: html_document
---
```{r}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r}
library(tidyverse)         # for reading in data, graphing, and cleaning
library(tidymodels)        # for modeling ... tidily
library(usemodels)         # for suggesting step_XXX() functions
library(glmnet)            # for regularized regression, including LASSO
library(naniar)            # for examining missing values (NAs)
library(lubridate)         # for date manipulation
library(moderndive)        # for King County housing data
library(vip)               # for variable importance plots
library(rmarkdown)         # for paged tables
library(readr)
library(fastDummies)
library(kableExtra)
library(DALEX)             # moDel Agnostic Language for Exploration and eXplanation (for model interpretation)  
library(DALEXtra)
theme_set(theme_minimal()) # my favorite ggplot2 theme :)
```
## Load the Data
```{r}
finalDATASET <- read_csv("FINALDATASET.csv")
macro_data <- read_csv("clean_macro - Sheet1.csv")
```

## Data Processing 
```{r}
final_data <- finalDATASET %>% 
  # make Sector dummy variables
  dummy_cols(select_columns = "Sector") %>% 
  # merge with updated macro data
  select(-CPALTT01USM657N_PC1,
         -GDP,
         -GDP_PC1,
         -M1SL_PC1 ,
         -M1SL,
         -PRICE,
         -Sell,
         -COMPANY) %>% 
  merge(macro_data) %>% 
  # convert Macro factors from characters to numeric
  mutate(across(c("CPALTT01USM657N_PC1","GDP","GDP_PC1","T10Y2Y","M1SL_PC1","M1SL"),
                as.numeric)) %>% 
  group_by(YEAR,Sector) %>% 
  # replacing the missing value with median od the industry in that year
  mutate(across(c(DEBTS,INVESTMENTS,CASH,VOLUME,EARNINGS,COGS,SALES,RECEIVABLE,INVENTORY),~replace(.,.==0,median(.)))) %>%
  # delete less important factors -> Exchange can possibily be deleted
  ungroup() %>%  
  mutate(across(c(!where(is.numeric),-"Name"),as.factor)) %>% 
  select(-Earnings_next_year,-observation_date) %>% 
  drop_na() 
```

```{r}
set.seed(327) #for reproducibility

data_split <- initial_split(final_data, 
                             prop = .75)
data_training <- training(data_split)
data_testing <- testing(data_split)

```

## Building Recipe
In the lasso model, to account for the fact the macroeconomic condition affects each industry differently, we also included the interaction term between GDP and sector dummies. 
```{r}
return_recipe <- recipe(PROFIT ~ ., #short-cut, . = all other vars
                       data = data_training) %>% 
  # filter to only have data after 2020
  step_filter(YEAR<2021) %>% 
  step_rm(Name,Sector,YEAR) %>% 
  #step_rm(GDP,M1SL,Name,Sector,YEAR) %>% 
  # add PE 
  step_mutate(PE = `MARKET CAP`/EARNINGS) %>%
  # Normalize all variables except for GDP
  step_normalize(all_predictors(), 
                 -all_nominal(),
                 -starts_with("Sector_")) %>% 
  # Create interaction terms
  step_interact(terms = ~c(GDP_PC1):starts_with("Sector_")) 
```

```{r}
return_recipe %>% 
  prep(data_training) %>%
  # using bake(new_data = NULL) gives same result as juice()
  # bake(new_data = NULL)
  juice() 
```

## Building the Lasso model

```{r}
return_linear_mod <- 
  # Define a lasso model 
  # I believe default is mixture = 1 so probably don't need 
  linear_reg(mixture = 1) %>% 
  # Set the engine to "glmnet" 
  set_engine("glmnet") %>% 
  # The parameters we will tune.
  set_args(penalty = tune()) %>% 
  # Use "regression"
  set_mode("regression")

```
```{r}
return_lm_wf <- 
  # Set up the workflow
  workflow() %>% 
  # Add the recipe
  add_recipe(return_recipe) %>% 
  # Add the modeling
  add_model(return_linear_mod)

penalty_grid <- grid_regular(penalty(),
                             levels = 10)

return_cv <- vfold_cv(data_training, v = 5)

return_lm_tune <- 
  return_lm_wf %>% 
  tune_grid(
    resamples = return_cv,
    grid = penalty_grid
    )

```
```{r}
return_lm_tune %>% 
  collect_metrics() %>% 
  filter(.metric == "rmse") %>% 
  ggplot(aes(x = penalty, y = mean)) +
  geom_point() +
  geom_line() +
  scale_x_log10(
   breaks = scales::trans_breaks("log10", function(x) 10^x),
   labels = scales::trans_format("log10",scales::math_format(10^.x))) +
  labs(x = "penalty", y = "rmse")
```
## Select tuning parameter
```{r}
best_param<-return_lm_tune %>% 
  select_best(metric = "rmse")

return_lasso_final_wf <- return_lm_wf %>% 
  finalize_workflow(best_param)

return_lasso_final_mod <- return_lasso_final_wf %>% 
  fit(data = data_training)
```
## Lasso results
The table below shows the estimate of each predictor in the Lasso Model.  We can see that indicators in macroeconomics are relatively important to predict the stock return. The change in inflation and GDP all remain significant after shrinking. The interaction terms between GDP and industries also showed importance, accounting for the fact that the macroeconomic condition affects each industry differently: it seems to affect the stock return in Communication Services, Energy, and Consumer Discretionary sectors less. 
```{r}
return_lasso_final_mod %>% 
  pull_workflow_fit() %>% 
  tidy() %>% 
  arrange(desc(estimate)) %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("striped", "bordered", "hover", "condensed")) %>% 
  scroll_box()
```
## Model Evaluation
### Prediciton precision 
The Lasso model does not perform well. It only explains 12% of the varaince in stock return and it also has a relatively high rmse. 
```{r}
return_lm_tune %>%
  select(id, .metrics) %>%
  unnest(.metrics) %>%
  filter(.metric == "rsq") %>% 
  filter(.config == "Preprocessor1_Model09") %>% 
  summarise(mean_rsq = mean(.estimate))

return_lm_tune %>%
  select(id, .metrics) %>%
  unnest(.metrics) %>%
  filter(.metric == "rmse") %>% 
  filter(.config == "Preprocessor1_Model09") %>% 
  summarise(mean_rmse = mean(.estimate))
# Originally 0.14 + 52
```

```{r eval = FALSE}
prediction <- predict(
  return_lasso_final_mod,
  new_data = data_training)

training_pred<-data_training %>% 
  mutate(.pred = prediction$.pred)
  
  
training_pred %>% 
  ggplot(aes(x = PROFIT, 
             y = .pred,
             color = YEAR)) +
  geom_point(alpha = .5, 
             size = .5) +
  geom_smooth(se = FALSE) +
  geom_abline(slope = 1, 
              intercept = 0, 
              color = "darkred") +
  geom_text(aes(label = Name),label.size = 0.15,data = training_pred %>% filter(PROFIT>400) )+
  labs(x = "Actual Return", 
       y = "Predicted Return") +
  scale_color_viridis_b()
```

### Overfitting
Below shows the r-square and rmse on testing data. They are very similar to those on trainning data so we would safely conclude that our model did not overfit. 
```{r}
return_lasso_test <- return_lasso_final_wf %>% 
  last_fit(data_split)

# Metrics for model applied to test data
return_lasso_test %>% 
  collect_metrics()
```
The following graphs 
```{r}
test_prediction <- predict(
  return_lasso_final_mod,
  new_data = data_testing)

testing_lasso_pred<-data_testing %>% 
  mutate(.pred = test_prediction$.pred)
  
  
testing_lasso_pred %>% 
  ggplot(aes(x = PROFIT, 
             y = .pred,
             color = YEAR)) +
  geom_point(alpha = .5, 
             size = .5) +
  geom_smooth(se = FALSE) +
  geom_abline(slope = 1, 
              intercept = 0, 
              color = "darkred") +
  geom_text(aes(label = Name),label.size = 0.15,data = testing_lasso_pred %>% filter(PROFIT>300) )+
  labs(x = "Actual Return", 
       y = "Predicted Return") + 
  scale_color_viridis_b()
```


## Interpretable Machine Learning 
Again, the boxplot and the histogram of residuals, the model does not predict the return well and this may because some extreme values. 
```{r}
lasso_explain <- 
  explain_tidymodels(
    model = return_lasso_final_mod,
    data = data_training %>% select(-PROFIT), 
    y = data_training %>%  pull(PROFIT),
    label = "lasso"
  )
```

```{r}
lasso_mod_perf <- model_performance(lasso_explain)
hist_plot <- 
  plot(lasso_mod_perf,
       geom = "histogram")
box_plot <-
  plot(lasso_mod_perf,
       geom = "boxplot")

hist_plot
box_plot

```
Below shows the feature importance plots generated by two different methods. We could see that macroeconomic indicators remained importance in both of the plots but market cap seems to be an crucial factor when we use permutation method.

```{r}
# Why year and sector are still in it??
set.seed(10) #since we are sampling & permuting, we set a seed so we can replicate the results
lasso_var_imp <- 
  model_parts(
    lasso_explain
    )

plot(lasso_var_imp, show_boxplots = TRUE)
```
```{r}
return_lasso_final_mod %>% 
  pull_workflow_fit() %>% 
  vip()
```




